#### This is a program for calculating the spearman correlations between cytokine and metabolite abundances.

### load the libraries

```{r}
### load the libraries
library(tidyverse)
library(psych)
library(ComplexHeatmap)
library(circlize)
library(ggpubr)
```

### load the data

```{r}
load("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/data/processed/T21_mixed_cyt_met.RData")
load("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/data/processed/D21_mixed_cyt_met.RData")

```


### Calculate the correlations within Cytokines

``` r
T21_cytokine_correlations <- T21.mixed$cytokines %>%
    corr.test(method = "spearman", adjust = "fdr")

D21_cytokine_correlations <- D21.mixed$cytokines %>%
    corr.test(method = "spearman", adjust = "fdr")

```

### Calculate the correlations within Metabolites

``` r
T21_metabolite_correlations <- T21.mixed$metabolites %>%
    corr.test(method = "spearman", adjust = "fdr")

D21_metabolite_correlations <- D21.mixed$metabolites %>%
    corr.test(method = "spearman", adjust = "fdr")

```

### Heatmaps

``` r
cyt_heatmap = Heatmap(T21_cytokine_correlations$
	r, 
	name = "correlation coefficient",
	row_dend_reorder = TRUE, 
	column_dend_reorder = TRUE, 
	show_row_dend = FALSE, 
	show_column_dend = TRUE, 
	column_names_gp = gpar(fontsize = 7),
	row_names_gp = gpar(fontsize = 7), 
	show_row_names = TRUE, 
	show_column_names = FALSE,
	column_split = 4,
	row_split = 4,
	row_title = NULL,
	column_title = NULL,
	col=colorRamp2(c(-.5, 0, .5), c("blue", "white", "red")),
	heatmap_legend_param = list(direction = "horizontal")
    )

pdf("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/correlation_heatmaps/T21_Cytokines_Correlation_Heatmap.pdf")
draw(cyt_heatmap, heatmap_legend_side = "bottom")
dev.off()

```

``` r
met_heatmap = Heatmap(T21_metabolite_correlations$r, 
	name = "correlation coefficient",
	row_dend_reorder = TRUE, 
	column_dend_reorder = TRUE, 
	show_row_dend = FALSE, 
	show_column_dend = TRUE, 
	column_names_gp = gpar(fontsize = 3),
	row_names_gp = gpar(fontsize = 3), 
	show_row_names = TRUE, 
	show_column_names =FALSE,
	column_split = 4,
    row_title = NULL,
	row_split = 4,
	column_title = NULL,
	col=colorRamp2(c(-.5, 0, .5), c("blue", "white", "red")),
	heatmap_legend_param = list(direction = "horizontal")
    )
pdf("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/correlation_heatmaps/T21_Metabolites_Correlation_Heatmap.pdf")
draw(met_heatmap, heatmap_legend_side = "bottom")
dev.off()

```

### Calculate the correlations across molecular classs

``` r
T21.omics <- cbind(T21.mixed$metabolites, T21.mixed$cytokines)

T21_correlations <- T21.omics %>%
    corr.test(method = "spearman", adjust = "fdr")

D21.omics <- cbind(D21.mixed$metabolites, D21.mixed$cytokines)

D21_correlations <- D21.omics %>%
    corr.test(method = "spearman", adjust = "fdr")
```


### Heatmap of multiomic correlations

``` r
all_heatmap = Heatmap(T21_correlations$r, 
	name = "correlation coefficient",
	row_dend_reorder = TRUE, 
	column_dend_reorder = TRUE, 
	show_row_dend = FALSE, 
	show_column_dend = TRUE, 
	column_names_gp = gpar(fontsize = 3),
	row_names_gp = gpar(fontsize = 2), 
	show_row_names = TRUE, 
	show_column_names =FALSE,
	column_split = 4,
    row_title = NULL,
	row_split = 4,
	column_title = NULL,
	col=colorRamp2(c(-.5, 0, .5), c("blue", "white", "red")),
	heatmap_legend_param = list(direction = "horizontal")
    )
pdf("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/correlation_heatmaps/T21_ALL_Correlation_Heatmap.pdf")
draw(all_heatmap, heatmap_legend_side = "bottom")
dev.off()

```



### Cytokine-metabolite scatter plots

#### Cytokines
``` r
T21_cytokine_correlations$long <- T21_cytokine_correlations$r %>%
    as.data.frame() %>%
    rownames_to_column("A") %>%
    pivot_longer(cols = -A, names_to = "B") %>%
    filter(A!=B)

T21_cytokine_correlations$p_long <- T21_cytokine_correlations$p %>%
    as.data.frame() %>%
    rownames_to_column("A") %>%
    pivot_longer(cols = -A, names_to = "B") %>%
    filter(A!=B)



T21_cytokine_correlations$long <- cbind(T21_cytokine_correlations$long, p = T21_cytokine_correlations$p_long, fdr = T21_cytokine_correlations$p.adj)

T21_cytokine_correlations$long <- T21_cytokine_correlations$long %>%
    filter(fdr < .1) %>%
    arrange(-value) 

options(ggpubr.parse_aes = FALSE)

sapply(1:10, function(i) {
    tmp.1 = T21_cytokine_correlations$long$A[i]
    tmp.2 = T21_cytokine_correlations$long$B[i]
    tmp.plot =  ggscatter(data = T21.mixed$cytokines,
                          x =  tmp.1,
                          y = tmp.2,
                          add = "reg.line",  # Add regression line
                          add.params = list(color = "blue", fill = "lightgray"), # Customize reg. line
                          conf.int = TRUE # Add confidence interval
                          ) +
        stat_cor( aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")), method = "spearman",label.sep = "\n", size = 8)+
        theme(axis.text = element_text(size =18),
              axis.title = element_text(size =20))
    pdf(paste0("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/correlation_scatter_plots/cytokines/",make.names(tmp.1), " - ", make.names(tmp.2), ".pdf"))
    print(tmp.plot)
    dev.off()
})

```



