#### This is a program for calculating the spearman correlations between cytokine and metabolite abundances.

### load the libraries

```{r}
### load the libraries
library(tidyverse)
library(ComplexHeatmap)
library(circlize)
source("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/R/cyt_met_correlation_functions.R")

```

### load the data

```{r}
load("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/data/processed/T21_mixed_cyt_met.RData")
load("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/data/processed/D21_mixed_cyt_met.RData")

```


### Calculate the correlations within Cytokines

``` r
T21_cytokine_correlations <- T21.mixed$cytokines %>%
    corr.test(method = "spearman", adjust = "fdr")

D21_cytokine_correlations <- D21.mixed$cytokines %>%
    corr.test(method = "spearman", adjust = "fdr")

```

### Calculate the correlations within Metabolites

``` r
T21_metabolite_correlations <- T21.mixed$metabolites %>%
    corr.test(method = "spearman", adjust = "fdr")

D21_metabolite_correlations <- D21.mixed$metabolites %>%
    corr.test(method = "spearman", adjust = "fdr")

```

### Heatmaps

``` r
cyt_heatmap = Heatmap(T21_cytokine_correlations$
	r, 
	name = "correlation coefficient",
	row_dend_reorder = TRUE, 
	column_dend_reorder = TRUE, 
	show_row_dend = FALSE, 
	show_column_dend = TRUE, 
	column_names_gp = gpar(fontsize = 7),
	row_names_gp = gpar(fontsize = 7), 
	show_row_names = TRUE, 
	show_column_names = FALSE,
	column_split = 4,
	row_split = 4,
	row_title = NULL,
	column_title = NULL,
	col=colorRamp2(c(-.5, 0, .5), c("blue", "white", "red")),
	heatmap_legend_param = list(direction = "horizontal")
    )

pdf("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/correlation_heatmaps/T21_Cytokines_Correlation_Heatmap.pdf")
draw(cyt_heatmap, heatmap_legend_side = "bottom")
dev.off()

```

``` r
met_heatmap = Heatmap(T21_metabolite_correlations$r, 
	name = "correlation coefficient",
	row_dend_reorder = TRUE, 
	column_dend_reorder = TRUE, 
	show_row_dend = FALSE, 
	show_column_dend = TRUE, 
	column_names_gp = gpar(fontsize = 3),
	row_names_gp = gpar(fontsize = 3), 
	show_row_names = TRUE, 
	show_column_names =FALSE,
	column_split = 4,
    row_title = NULL,
	row_split = 4,
	column_title = NULL,
	col=colorRamp2(c(-.5, 0, .5), c("blue", "white", "red")),
	heatmap_legend_param = list(direction = "horizontal")
    )
pdf("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/correlation_heatmaps/T21_Metabolites_Correlation_Heatmap.pdf")
draw(met_heatmap, heatmap_legend_side = "bottom")
dev.off()

```

### Calculate the correlations across molecular classs

``` r
T21.omics <- cbind(T21.mixed$metabolites, T21.mixed$cytokines)

T21_correlations <- T21.omics %>%
    corr.test(method = "spearman", adjust = "fdr")

D21.omics <- cbind(D21.mixed$metabolites, D21.mixed$cytokines)

D21_correlations <- D21.omics %>%
    corr.test(method = "spearman", adjust = "fdr")
```


### Heatmap of multiomic correlations

``` r
all_heatmap = Heatmap(T21_correlations$r, 
	name = "correlation coefficient",
	row_dend_reorder = TRUE, 
	column_dend_reorder = TRUE, 
	show_row_dend = FALSE, 
	show_column_dend = TRUE, 
	column_names_gp = gpar(fontsize = 3),
	row_names_gp = gpar(fontsize = 2), 
	show_row_names = TRUE, 
	show_column_names =FALSE,
	column_split = 4,
    row_title = NULL,
	row_split = 4,
	column_title = NULL,
	col=colorRamp2(c(-.5, 0, .5), c("blue", "white", "red")),
	heatmap_legend_param = list(direction = "horizontal")
    )
pdf("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/correlation_heatmaps/T21_ALL_Correlation_Heatmap.pdf")
draw(all_heatmap, heatmap_legend_side = "bottom")
dev.off()

```



### Cytokine-metabolite scatter plots


``` r


T21_cytokine_correlations <- LongCorr(T21_cytokine_correlations)
T21_metabolite_correlations <- LongCorr(T21_metabolite_correlations)
T21_correlations <- LongCorr(T21_correlations)
T21_correlations$long <- T21_correlations$long %>%
    filter(A %in% names(T21.mixed$cytokines) & B %in% names(T21.mixed$metabolites))

corrScatterPlotsSingle(psych_object = T21_cytokine_correlations,
                       molecule_class = "cytokines",
                       omics_data = T21.mixed)


corrScatterPlotsSingle(psych_object = T21_metabolite_correlations,
                       molecule_class = "metabolites",
                       omics_data = T21.mixed)



corrScatterPlotsMultiple(psych_object = T21_correlations,
                       molecule_classes = c("cytokines","metabolites"),
                       omics_data = T21.mixed)


```



