---
title: "Pathway Annotation"
author: "Lucas Gillenwater"
date: "`r Sys.Date()`"
output: html_document
---
For downstream enrichment analyses, we annotated features based on prior knowledge of biological pathways. 

# Necessary Libraries
```{r}
library(biomaRt)
library(data.table)
```
# Load the molecular data

```{r}
load("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/data/processed/T21_mixed_cyt_met.Rdata")

```

# Metabolites
# First create a list containing a vector of the compounds to be queried (separated by a semi-colon)  
# and another character vector containing the compound id type.
# The items in the list MUST be queryList and inputType
# Valid input types are: "name", "hmdb", "kegg", "pubchem", "chebi", "metlin"

``` r
mets = names(T21.mixed$metabolites)
name.vec <- paste(mets, collapse = ";")
toSend = list(queryList = name.vec, inputType = "name")

library(httr)
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("metaboliteIDmapping", force = TRUE)

library( metaboliteIDmapping)

# The MetaboAnalyst API url
call <- "http://api.xialab.ca/mapcompounds"

# Use httr::POST to send the request to the MetaboAnalyst API
# The response will be saved in query_results
query_results <- httr::POST(call, body = toSend, encode = "json")

# Check if response is ok (TRUE)
# 200 is ok! 401 means an error has occured on the user's end.
query_results$status_code==200

# Parse the response into a table
# Will show mapping to "hmdb_id", "kegg_id", "pubchem_id", "chebi_id", "metlin_id", "smiles" 
query_results_text <- content(query_results, "text", encoding = "UTF-8")
query_results_json <- rjson::fromJSON(query_results_text)
query_results_list = lapply(query_results_json, as.character)
query_results_table = do.call(cbind, query_results_list)

write.csv(met_anno, file = "~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/data/pathway/metabolite_annotations.csv", row.names = F)


### After manual curation read back in annotated results
met_anno = read.csv("./data/pathway/metabolite_annotations_all.csv")

to_pathway = character()
for(i in 1:nrow(met_anno)){
	tmp = ifelse(!is.na(met_anno[i,"CAS"]), sprintf("CAS:%s",met_anno[i,"CAS"] ),
			ifelse(!is.na(met_anno[i,"KEGG"]), sprintf("kegg:%s",met_anno[i,"KEGG"]),
			ifelse(!is.na(met_anno[i,"ChEBI"]), sprintf("chebi:%s",met_anno[i,"ChEBI"] ),
		    ifelse(!is.na(met_anno[i,"HMDB"]), sprintf("hmdb:%s",met_anno[i,"HMDB"] ),
			sprintf("LIPIDMAPS:%s",met_anno[i,"Lipid.Maps"] )))))
	to_pathway = c(to_pathway, tmp)
}


install.packages("remotes")
remotes::install_github("Mathelab/RaMP-DB")
library(RaMP)
pkg.globals <- setConnectionToRaMP(dbname="ramp2",username="root",conpass="",
                                   host = "localhost")
test = chemicalClassSurvey(mets = to_pathway)
test = as.data.frame(test$met_classes)
test = test[grep("sub_class", test$class_level_name),]

pathway_map = as.data.frame(cbind(metabolite = met_anno$Query, to_pathway))

met_classes = merge(pathway_map, test, by.x = "to_pathway", by.y = "sourceId", all = T )
met_classes = dcast(met_classes, metabolite ~class_name)
met_classes = t(met_classes)
met_classes = apply(met_classes,2, function(x){
					if(!is.na(any(x == "lipidmaps"))){
						ifelse(x == "hmdb", NA, x)
					} else {
						x = x 
					}
				})


colnames(met_classes) = met_classes[1,]
met_classes =met_classes[-1,]
met_classes = ifelse(!is.na(met_classes),1,0)
write.csv(met_classes, "./data/pathway/metabolite_pathway.csv")
save(met_classes, file = "./data/pathway/metabolite_pathway.RData")





# Cytokines
## Getting Ensembl IDs
Annotation begins by matching the cytokine names to proteins in the protein atlas to get the Ensembl ids. Since mot all of the cytokine names had a perfect match, I had to hand annotate the remaining features. 
```{r}
cytokines = rownames(all.mixed$cytokines)

# download the human protein atlas
dir.create("./data/pathway_annotation")
system("wget https://www.proteinatlas.org/download/proteinatlas.tsv.zip  -P ./data/pathway_annotation")
system("unzip ./data/pathway_annotation/proteinatlas.tsv.zip -d ./data/pathway_annotation")

protein_atlas <- read.delim("./data/pathway_annotation/proteinatlas.tsv")

protein_atlas_cytokine_id <- sapply(cytokines, function(x) {which(grepl(x, protein_atlas$Gene, ignore.case = T) | grepl(x, protein_atlas$Gene.synonym, ignore.case = T))})
protein_atlas_cytokine_id <- unlist(protein_atlas_cytokine_id)

protein_atlas_cytokine <- protein_atlas[protein_atlas_cytokine_id,1:5]

write.csv(protein_atlas_cytokine,"./data/pathway_annotation/cytokine_annotation_raw.csv")
## I proceeded to manually annotate the remaining cytokines by manual search in the proteinatlas tsv file. The final file was saved as cytokine_annotation_processed.csv

```
## Matching the ensembl ids to GO processes
```{r}
cytokine_annotation = read.csv("./data/pathway_annotation/cytokine_annotation_processed.csv")
cytokine_annotation = cbind(cytokine = cytokines, cytokine_anno)

cyt_ens <- cytokine_anno$Ensembl

ensembl <- useEnsembl(biomart = "genes", dataset = "hsapiens_gene_ensembl")
resultTable <- biomaRt::getBM(attributes = c("ensembl_gene_id", "name_1006", "go_linkage_type", "namespace_1003"),       
                      filters    = "ensembl_gene_id",       
                      values     = cyt_ens,         
                      mart       = ensembl) 
resultTable = resultTable[resultTable$go_linkage_type == "IDA",]
resultTable = resultTable[resultTable$namespace_1003 == "biological_process",]

cytokine_go <- merge(resultTable, cytokine_annotation, by.x = "ensembl_gene_id", by.y = "Ensembl", all.y = TRUE)

cyt_go <- table(cytokine_go$name_1006, cytokine_go$cytokine )



cyt_path = t(cyt_go)
ind = !duplicated(cyt_path[,1:ncol(cyt_path)])
cyt_path = cyt_path[, ind]
cyt_path = melt(cyt_path)
colnames(cyt_path)[1:2] = c("X", "pathway")
cyt_path$value = as.numeric(cyt_path$value)

save(cyt_path, file = "./data/pathway/cytokine_pathway.RData")


```


