# Partial Correlation Visualization

This script utilized the results from the partial correlation analysis to create figures for publication. 
```{r loading data}
load("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/partial_correlation_results/partial_cor_gsea_results_1_15_24.RData")
load("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/partial_correlation_results/full_partial_correlation_results_1_22_24.RData")
#transcription_factors <- read.delim("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/data/transcription_factors.txt", header = F)
load("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/partial_correlation_results/full_z_percent_1_15_24.RData")
load("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/partial_correlation_results/long_z_1_15_24.RData") 
load("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/partial_correlation_results/partial_corrlation_gsea_results_formatted_1_15_24.RData") 
load("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/partial_correlation_results/difference_in_clusters.RData")
names(long_z) <-as.character(c("T21","D21",1,2,3,4))
load(file = "~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/partial_correlation_results/cluster_correlations.RData")
source("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/R/partial_correlation_functions.R")
pathways_all <- gmtPathways("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/data/pathway/h.all.v2023.1.Hs.symbols.gmt")
met_path <- get(load("/Users/lucas/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/data/pathway/metabolite_pathway.RData"))

```

## data prep
```{r  data}
library(tidyverse)
toplot_clusters <- lapply(gsea_results_formatted, function(x) {
  x[,colSums(is.na(gsea_results_formatted[[1]])) <1]
})
toplot_clusters <- lapply(toplot_clusters, function(x) {
  x %>%
    dplyr::select(contains("NES"))
}) 
for(x in 2:6) {
  names(toplot_clusters[[x]]) <- names(toplot_clusters[[1]])
}
names(toplot_clusters) <-c("T21", "D21", 1,2,3,4)


### Need to name the gsea result data frames
for(i in 2:6) {
    names(gsea_results_formatted[[i]]) <- names(gsea_results_formatted[[1]])
}

names(gsea_results_formatted) <- c("T21", "D21", 1,2,3,4)
```

#### Figure 2A
```{r percent difference plots}
library(tidyverse)

tmp_toplot <- full_results %>%
  filter(cluster == "T21") %>%
  filter(gene == "IDO1") %>%
  filter(cytokine ==  "IFN-gamma")


long_z[[1]] %>%
         filter(cyt_met %in% c("IFN-gamma-kynurenine", "IFN-gamma-acyl-C12 (O-dodecanoyl-carnitine)" )) %>%
  filter(gene == "IDO1")%>%
  dplyr::select(cytokine, metabolite, gene, cluster, cor_cyt_met_r, partial_cor_cyt_met_r) %>%
  pivot_longer(cols = -c(cytokine,metabolite, gene,cluster)) %>%
  mutate(name = c("direct correlation", "partial correlation"))

p1 <- ggplot(tmp_toplot, aes(x = as.factor(name), y = value, fill = name )) +
  geom_bar(stat = "identity") +
  theme_classic() +
  ylab("correlation coefficient") +
  theme(legend.position = "none") +
  xlab("")

pdf(paste0("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/partial_correlation_results/figures/percent_change.pdf"), width = 3, height = 3)
print(p1)
dev.off()
```

#### Figure 2B
```{r}
library(tidyverse)
library(ggplot2)
```

``` {r}
key_cyt_met <- "IFN-gamma-kynurenine"

single_rel_z <- long_z[[1]] %>%
  filter(`cyt_met` == key_cyt_met) %>%
  as.data.frame() %>%
  arrange(-combined_Z) %>%
  mutate(rank = row_number())



p1 <- ggplot(single_rel_z, aes(x = rank, y = combined_Z)) +
  geom_point() +
  geom_point(data = single_rel_z[single_rel_z$gene == "IDO1",], aes(x = rank, y = combined_Z),  color ="red")+
  annotate("text", y= single_rel_z[single_rel_z$gene == "IDO1","combined_Z"],x =single_rel_z[single_rel_z$gene == "IDO1","rank"]+ 1000, angle = 0,  label="IDO1", color = "red", size = 6) +
  geom_point(data = single_rel_z[single_rel_z$gene == "STAT1",], aes(x = rank, y = combined_Z),  color ="darkgreen")+
  annotate("text", y= single_rel_z[single_rel_z$gene == "STAT1","combined_Z"],x =single_rel_z[single_rel_z$gene == "STAT1","rank"]+ 1050, angle = 0,  label="STAT1", color = "darkgreen", size = 6)+
  theme_classic() +
  xlab("rank") +
  ylab("Z score")

pdf(paste0("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/partial_correlation_results/figures/",key_cyt_met,"_Z_trend_plot.pdf"), width = 10, height = 8)
print(p1)
dev.off()



```

### Figure 2C

```{r}
### placeholder for long_z file
key_gene <- "STAT1"

single_rel_z <- long_z[[1]] %>%
  filter(gene == key_gene) %>%
  as.data.frame() %>%
  arrange(-combined_Z) %>%
 mutate(rank = row_number())

p1 <- ggplot(single_rel_z, aes(x = rank, y = combined_Z)) +
  geom_point() +
  geom_point(data = single_rel_z[single_rel_z$cyt_met == "IFN-gamma-kynurenine",], aes(x = rank, y = combined_Z),  color ="red")+
    annotate("text", y= single_rel_z[single_rel_z$cyt_met == "IFN-gamma-kynurenine","combined_Z"],x =single_rel_z[single_rel_z$cyt_met == "IFN-gamma-kynurenine","rank"]+ 100, angle = 0,  label="IFN-gamma-kynurenine", color = "red", size = 6) +
  theme_classic() +
  xlab("rank") +
  ylab("Z score")

pdf(paste0("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/partial_correlation_results/figures/",key_gene,"_Z_trend_plot.pdf"), width = 10, height = 8)
print(p1)
dev.off()
```


## Figure 2D

## Heatmaps filtered by genes in geneset and ordered by gene set enrichment

``` r
T21_filtered_full <- T21_filtered %>%
    filter(cor_cyt_met_p < .05)

DS_pathways <- c("HALLMARK_INTERFERON_GAMMA_RESPONSE", 
                 "HALLMARK_INTERFERON_ALPHA_RESPONSE",
                 "HALLMARK_HEME_METABOLISM", 
                 "HALLMARK_INFLAMMATORY_RESPONSE",
                 "HALLMARK_OXIDATIVE_PHOSPHORYLATION",
                 "HALLMARK_MTORC1_SIGNALING",
                 "HALLMARK_G2M_CHECKPOINT",
                 "HALLMARK_E2F_TARGETS")

DS_pathways <- DS_pathways[c(1,3,5)]
unique_pathways <- sort(unique(met_path$Pathway))
n <- length(unique_pathways)
qual_col_pals <- brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector <- unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))
col_vector <- col_vector[1:n]
names(col_vector) <- unique_pathways
col_vector <- c(col_vector, "unknown" = "black")

toplot1 <- geneSetUnion(key_pathways = DS_pathways,
             pathways_all = pathways_all,
             met_path = met_path,
             formatted_gsea = gsea_results_formatted[[1]],
             z_scores = long_z[[1]],
             num_cyt_met = 400
             )

```


## Exploring top ranked cytokine-metabolite relationships and genes in a particular pathway
``` r
## IFN_gamma

IFN_IFN_cyt_mets <- OrderedGSEA(
               formatted_gsea = gsea_results_formatted[['T21']],
    key_pathway = "HALLMARK_INTERFERON_GAMMA_RESPONSE",
    p_thresh = .05)

dim(IFN_cyt_mets)
head(IFN_cyt_mets$cyt_met,20)

cytokines <- unique(T21_filtered_full$cytokine)
sig_mets <- as.character(IFN_cyt_mets$cyt_met)

sig_mets <- unique(unlist(sapply(cytokines, function(x) {
    tmp <- gsub(paste0(x,"-"), "",sig_mets[grepl(x, sig_mets)])
    return(tmp)
})))

source("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/r/enrichment_functions.R")
metEnrich(sig_mets, met_path)

met_path %>% filter(Pathway %in% c("Other sugars", "Saturated Fatty acids")) %>% select(compound)




## Heme Metabolism
heme_heme_cyt_mets <- OrderedGSEA(
               formatted_gsea = gsea_results_formatted[['T21']],
    key_pathway = "HALLMARK_HEME_METABOLISM",
    p_thresh = .05)

dim(heme_heme_cyt_mets)
head(heme_heme_cyt_mets,20)

met_path %>% select(name, Pathway)
sig_mets <- as.character(heme_cyt_mets$cyt_met)
sig_mets <- unique(unlist(sapply(cytokines, function(x) {
    tmp <- gsub(paste0(x,"-"), "",sig_mets[grepl(x, sig_mets)])
    return(tmp)
})))
metEnrich(sig_mets, met_path)




## Oxidative Phosphorylation
op_op_cyt_mets <- OrderedGSEA(
               formatted_gsea = gsea_results_formatted[['T21']],
    key_pathway = "HALLMARK_OXIDATIVE_PHOSPHORYLATION",
    p_thresh = .05)

dim(op_cyt_mets)
head(op_cyt_mets,20)

sig_mets <- as.character(op_cyt_mets$cyt_met)
sig_mets <- unique(unlist(sapply(cytokines, function(x) {
    tmp <- gsub(paste0(x,"-"), "",sig_mets[grepl(x, sig_mets)])
    return(tmp)
})))
metEnrich(sig_mets, met_path)




```


#### Venn Diagram of mediation signatures 
```{r}
DS_pathways <- c("HALLMARK_INTERFERON_GAMMA_RESPONSE", 
                 "HALLMARK_HEME_METABOLISM", 
                 "HALLMARK_OXIDATIVE_PHOSPHORYLATION"
                 )

## if (!require(devtools)) install.packages("devtools")
## devtools::install_github("yanlinlin82/ggvenn")

library(ggvenn)

x <- list(IFN_cyt_mets$cyt_met, heme_cyt_mets$cyt_met, op_cyt_mets$cyt_met)
names(x) <- gsub("HALLMARK_", "", DS_pathways)

venn <- ggvenn(
  x, 
  fill_color = c("#0073C2FF", "#EFC000FF", "#868686FF", "#CD534CFF"),
  stroke_size = 0.5, set_name_size = 3
  )


ggsave(filename = "~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/venn_diagrams/mediation_signatures_venn.pdf", width = 10, height =5, units ="in")


IFN_heme <- intersect(x[[1]],x[[2]])
write.csv(as.data.frame(IFN_heme), file = "~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/tables/IFN_heme_overlap.csv", row.names = F, quote = F)

IFN_op <- intersect(x[[1]],x[[3]])
write.csv(as.data.frame(IFN_op), file = "~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/tables/IFN_op_overlap.csv", row.names = F, quote = F)

heme_op <- intersect(x[[2]],x[[3]])
write.csv(as.data.frame(heme_op), file = "~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/tables/heme_op_overlap.csv", row.names = F, quote = F)
```


``` r

source("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/R/partial_correlation_functions.R")
multiHeatmapOrderedGSEA(toplot1 = toplot1, 
                        toplot2 = toplot_clusters[[1]],
                        key_pathways = DS_pathways,
                        pathways_all = pathways_all,
                        met_path = met_path,
                        formatted_gsea = gsea_results_formatted[[1]] ,
                        full_results = T21_filtered_full,
                        output_file = "~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/mediation_heatmaps/T21_heatmaps/2D_filtered_ordered-IFN-gamma.pdf",
                        order_pathway = "HALLMARK_INTERFERON_GAMMA_RESPONSE",
                        col_vector = col_vector)



multiHeatmapOrderedGSEA(toplot1 = toplot1,
                        toplot2 = toplot_clusters[[1]],
                        key_pathways = DS_pathways,
                        pathways_all = pathways_all,
                        met_path = met_path,
                        formatted_gsea = gsea_results_formatted[[1]] ,
                        full_results = T21_filtered_full,
                        output_file = "~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/mediation_heatmaps/T21_heatmaps/2D_filtered_ordered-HEME-METABOLISM.pdf",
                        order_pathway = "HALLMARK_HEME_METABOLISM",
                        col_vector = col_vector)
multiHeatmapOrderedGSEA(toplot1 = toplot1,
                        toplot2 = toplot_clusters[[1]],
                        key_pathways = DS_pathways,
                        pathways_all = pathways_all,
                        met_path = met_path,
                        formatted_gsea = gsea_results_formatted[[1]] ,
                        full_results = T21_filtered_full,
                        output_file = "~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/mediation_heatmaps/T21_heatmaps/2D_filtered_ordered-OXIDATIVE-PHOSPHORYLATION.pdf",
                        order_pathway = "HALLMARK_OXIDATIVE_PHOSPHORYLATION",
                        col_vector = col_vector)

```



### Heatmaps by cluster
```r
source("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/R/partial_correlation_functions.R")
names(long_z) <-as.character(c("T21","D21",1,2,3,4))

load(file = "~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/partial_correlation_results/cluster_correlations.RData")
all_correlations <- lapply(all_correlations, function(x) {
    x %>% mutate(cyt_met = paste0(A, "-", B))
})

cyt_mets <- filterCyt_Mets(toplot = toplot1,
               formatted_gsea = gsea_results_formatted[[1]],
               key_pathway = "HALLMARK_INTERFERON_GAMMA_RESPONSE",
               gene_sets = pathways_all,
               num_cyt_met = 15)

cluster_ranks <- clusterGeneRanks(cyt_mets = cyt_mets,
                                  long_z_list = long_z,
                                  key_pathway = "HALLMARK_INTERFERON_GAMMA_RESPONSE",
                                  gene_sets = pathways_all)

names(cluster_ranks) <- names(long_z)
cluster_heatmaps <- clusterHeatmapWrapper(cluster_ranks, name = "Interferon Gamma Response",all_correlations, output_file = "~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/mediation_heatmaps/cluster_heatmaps/IFN-gamma.pdf")


cyt_mets <- filterCyt_Mets(toplot = toplot1,
               formatted_gsea = gsea_results_formatted[[1]],
               key_pathway = "HALLMARK_HEME_METABOLISM",
               gene_sets = pathways_all,
               num_cyt_met = 15)
cluster_ranks <- clusterGeneRanks(cyt_mets = cyt_mets,
                                  long_z_list = long_z,
                                  key_pathway = "HALLMARK_HEME_METABOLISM",
                                  gene_sets = pathways_all)
names(cluster_ranks) <- names(long_z)
cluster_heatmaps <- clusterHeatmapWrapper(cluster_ranks, name = "Heme Metabolism",all_correlations, output_file = "~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/mediation_heatmaps/cluster_heatmaps/Heme.pdf")



cyt_mets <- filterCyt_Mets(toplot = toplot1,
               formatted_gsea = gsea_results_formatted[[1]],
               key_pathway = "HALLMARK_OXIDATIVE_PHOSPHORYLATION",
               gene_sets = pathways_all,
               num_cyt_met = 15)
cluster_ranks <- clusterGeneRanks(cyt_mets = cyt_mets,
                                  long_z_list = long_z,
                                  key_pathway = "HALLMARK_OXIDATIVE_PHOSPHORYLATION",
                                  gene_sets = pathways_all)

names(cluster_ranks) <- names(long_z)
cluster_heatmaps <- clusterHeatmapWrapper(cluster_ranks, name = "Oxidative Phosphorylation",all_correlations, output_file = "~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/mediation_heatmaps/cluster_heatmaps/ox-phos.pdf")
```



### Heatmaps by cluster filtered by cyt_mets with significant differences in correlation 

``` r
source("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/R/partial_correlation_functions.R")

all_correlations <- lapply(all_correlations, function(x) {
    x %>% mutate(cyt_met = paste0(A, "-", B))
})

corr_diff <- lapply(corr_diff, function(x) {
    x %>% mutate(cyt_met = paste0(A, "-", B))
})

overall_diff <- lapply (corr_diff[c(1:4, 6)], function(x) {
    x %>% top_n(12,diff) %>% .$cyt_met
})



cyt_mets <- filterCyt_Mets(toplot = toplot1,
               formatted_gsea = gsea_results_formatted[[1]],
               key_pathway = "HALLMARK_INTERFERON_GAMMA_RESPONSE",
               gene_sets = pathways_all,
               num_cyt_met = 634)


pathway_diff <- lapply(overall_diff, function(x) {
    intersect(x,cyt_mets)
})
top_diff <- unlist(pathway_diff[c("D21", 1,2,3,4,5)])


diff_cyt_mets <- intersect( top_diff, cyt_mets)
diff_cyt_mets
cluster_ranks <- clusterGeneRanks(cyt_mets = diff_cyt_mets,
                                  long_z_list = long_z,
                                  key_pathway = "HALLMARK_INTERFERON_GAMMA_RESPONSE",
                                  gene_sets = pathways_all)
names(cluster_ranks) <- names(long_z)


cluster_heatmaps <- clusterHeatmapWrapper(cluster_ranks, name = "Interferon Gamma Response",all_correlations, output_file = "~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/mediation_heatmaps/cluster_heatmaps/IFN-gamma_diff_filtered.pdf")
```


### Global cluster enrichment by pathways
Enrichment over pathway between clusters
```r
cluster_enrichment <- lapply(names(pathways_all), function(x) clusterPathwayEnrichment(x,gsea_results))
names(cluster_enrichment) <- names(pathways_all)
cluster_enrichment <- lapply(cluster_enrichment, unlist)
cluster_enrichment <- cluster_enrichment %>% bind_rows()
cluster_enrichment <- as.data.frame(cluster_enrichment)
rownames(cluster_enrichment) <- names(pathways_all)
cluster_enrichment <- cluster_enrichment[-nrow(cluster_enrichment),]
rownames(cluster_enrichment) <- gsub("HALLMARK", "", rownames(cluster_enrichment))
rownames(cluster_enrichment) <- gsub("_", " ", rownames(cluster_enrichment))
normalized_enrichment <- cluster_enrichment[,3:6]/cluster_enrichment[,1]

col_fun <- colorRamp2(c(0.7,1,1.3), c("blue","white", "red"))
p1 <- Heatmap(as.matrix(normalized_enrichment),
              name = "T21-normalized\nNES",
              column_order= names(normalized_enrichment),
              show_row_dend = F,
              col = col_fun,
              row_names_side = "left",
              row_names_gp = gpar(fontsize = 8),
              column_names_side = "top"
              )
pdf("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/mediation_heatmaps/global_pathway_heatmaps/T21_normalized_NES_global_pathway_heatmap.pdf")
print(p1)
dev.off()


normalized_enrichment <- cluster_enrichment[,c(1,3:6)]/cluster_enrichment[,2]
col_fun <- colorRamp2(c(0.7,1,1.3), c("blue","white", "red"))
p1 <- Heatmap(as.matrix(normalized_enrichment),
              name = "D21-normalized\nNES",
              column_order= names(normalized_enrichment),
              show_row_dend = F,
              col = col_fun,
              row_names_side = "left",
              row_names_gp = gpar(fontsize = 8),
              column_names_side = "top"
              )
pdf("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/mediation_heatmaps/global_pathway_heatmaps/D21_normalized_NES_global_pathway_heatmap.pdf")
print(p1)
dev.off()
```

### Difference of enrichment of cytokine-metabolites in cluster vs T21
This plot shows the difference in enrichment scores for a pathway between the cluster and individuals with T21
```r


names(gsea_results) <- c("T21", "D21", 1,2,3,4)

clusterDiffPlot(pathway = "HALLMARK_INTERFERON_GAMMA_RESPONSE",
                key_cluster = 3,
                num_cyt_met = 20)

clusterDiffPlot(pathway = "HALLMARK_INTERFERON_GAMMA_RESPONSE",
                key_cluster = 1,
                num_cyt_met = 20)


```


### Enrichment of metabolic classes mediated

``` r
names(gsea_results) <- c("T21", "D21", 1,2,3,4)

cluster_gsea <- clusterCytMetNES(key_pathway = "HALLMARK_INTERFERON_GAMMA_RESPONSE",
                 cluster = 1,
                 gsea_results = gsea_results,
                 num_cyt_met = 634)

cytokines <- unique(T21_filtered_full$cytokine)
sig_mets <- as.character(unique(cluster_gsea$cyt_met))

sig_mets <- unique(unlist(sapply(cytokines, function(x) {
    tmp <- gsub(paste0(x,"-"), "",sig_mets[grepl(x, sig_mets)])
    return(tmp)
})))

source("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/r/enrichment_functions.R")
metEnrich(sig_mets, met_path)

cluster_gsea <- clusterCytMetNES(key_pathway = "HALLMARK_INTERFERON_GAMMA_RESPONSE",
                 cluster = 3,
                 gsea_results = gsea_results,
                 num_cyt_met = 634)

cytokines <- unique(T21_filtered_full$cytokine)
sig_mets <- as.character(unique(cluster_gsea$cyt_met))

sig_mets <- unique(unlist(sapply(cytokines, function(x) {
    tmp <- gsub(paste0(x,"-"), "",sig_mets[grepl(x, sig_mets)])
    return(tmp)
})))

source("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/r/enrichment_functions.R")
metEnrich(sig_mets, met_path)

```




### Gene centric differences in Z score by cluster

```r


scores <- clusterGeneZscores(key_gene= "STAT1",
                             long_z_list = long_z,
                             cyt_mets = unique(cluster_gsea$cyt_met))

col_fun <- colorRamp2(c(-15,0,15), c("blue","white", "red"))
p1 <- Heatmap(t(scores),
        name = "Z score",
        column_order= as.character(unique(cluster_gsea$cyt_met)),
        row_order = colnames(scores),
        show_row_dend = F,
        row_names_side = "left",
        column_names_gp = gpar(fontsize = 8),
        column_names_rot = 80,
        column_names_side = "bottom",
        column_title = "STAT1",
        col = col_fun
        )
pdf("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/mediation_heatmaps/Z_score_heatmaps/STAT1.pdf",
    height = 6,
    width = 10)
p1
dev.off()

```






### Heatmap of top mediators across cyt-met relationships

```r
cluster_gsea <- clusterCytMetNES(key_pathway = "HALLMARK_INTERFERON_GAMMA_RESPONSE",
                 cluster = 3,
                 gsea_results = gsea_results,
                 num_cyt_met = 20)
scores <- topMediatorsCluster(long_z_list = long_z,
                              cyt_mets = unique(cyt_mets),
                              cluster = '3')

ha = HeatmapAnnotation(gene = anno_text(scores$gene, gp = gpar(fontsize = 8)))

toplot <- t(scores[,2:7])
toplot <- apply(toplot,2,as.numeric)
rownames(toplot) <- names(scores)[2:7]

col_fun <- colorRamp2(c(-15,0,15), c("blue","white", "red"))
p1 <- Heatmap(toplot,
        name = "Z score",
        column_order= colnames(toplot),
        row_order = rownames(toplot),
        show_row_dend = F,
        row_names_side = "left",
        column_names_gp = gpar(fontsize = 8),
        column_names_rot = 80,
        column_names_side = "bottom",
        top_annotation = ha,
        col = col_fun,
        heatmap_legend_param = list(legend_direction = "horizontal")
        )

pdf("~/OneDrive - The University of Colorado Denver/Projects/ImmunoMetabolicSubtypes/results/mediation_heatmaps/Z_score_heatmaps/IFN-gamma-cyt-mets.pdf",
    height = 6,
    width = 12)
draw(p1, heatmap_legend_side = "bottom")
dev.off()

```



